\begin{abstract}

\dga{Don't make it cloudy for EuroSys!  Instead discuss the traditional tradeoff
between heavyweightness/lightweightness and cooperativity/preemptiveness.}

\mk{We provide the very first (weaken this) preemptible functions.  And on this, we
build the first blah blah preemptive userland thread library.}
    
In a world where programs increasingly execute on rented datacenter computing
resources, CPU time is money.  Both application developers and cloud providers would
benefit from an easy abstraction for enforcing execution time limits at sub-process
granularity.  We propose one such abstraction that is \textbf{lightweight} with
respect to the OS and scheduler, \textbf{preemptive} to guard against tight
computation loops, and uses an interface featuring structured control flow centered
around familiar \textbf{functions}.  We present \textit{libinger}, a
language-independent prototype implemented entirely in userland that functions atop
the existing system stack with no major changes, and requires only light runtime
support.  We then assess the overheads of this runtime support.  Finally, we present
a few case studies demonstrating possible uses of the abstraction and discuss its
end-to-end performance.  In particular, we show how lightweight preemptible functions
can be trivially applied to implement a preemptive thread library in userland.

\solb{``light runtime support'': Intro or design should explain that we require neither
garbage collection nor kernel changes}

\end{abstract}
