\section{Timed functions: \textit{libinger}}
\label{sec:libinger}

To address the literature's shortcomings, we have developed
\textit{libinger},\footnote{In the style of GNU's \textit{libiberty}, we named our
system for the command-line switch used to link against it.  As the proverb goes,
``Don't want your function calls to linger?  Link with \texttt{-linger}.''}
a library providing a small API for timed function dispatch
(Listing~\ref{lst:ingerapi}):
\begin{itemize}
\item \texttt{launch()} invokes an ordinary function \texttt{func} with a
time cap of \texttt{time\_us}.  The call to \texttt{launch()} returns when
\texttt{func}
completes, or after approximately \texttt{time\_us} microseconds if \texttt{func} has
not returned
by then.  In the latter case, \textit{libinger} returns an opaque continuation
object recording the execution state.
\item \texttt{resume()} causes a preemptible function to continue after a timeout.
If execution again times out, \texttt{resume()} updates its continuation so the
process may be repeated.  Resuming a function that has already returned has no
effect.
\end{itemize}

\begin{figure}
\begin{lstlisting}[label=lst:ingerapi,caption=Preemptible functions core interface]
struct linger_t {
	bool is_complete;
	cont_t continuation;
};

linger_t launch(Function func, u64 time_us, void *args);
void resume(linger_t *cont, u64 time_us);
\end{lstlisting}
\begin{lstlisting}[label=lst:use, caption=Preemptible function usage example]
linger = launch(task, TIMEOUT, NULL);
if (!linger.is_complete) {
	// Save `linger` to a task queue to resume later
	task_queue.push(linger);
}

// Handle other tasks
...

// Resume `task` at some later point
linger = task_queue.pop();
resume(&linger, TIMEOUT);
\end{lstlisting}
\end{figure}

Listing~\ref{lst:use} shows an example use of \textit{libinger}
in a task queue manager designed to prevent latency-critical tasks from blocking
behind longer-running
ones. The caller invokes a task with a timeout. If the task does not complete
within the allotted time, the caller saves its continuation in the task queue,
handles other tasks, and later resumes the first task.
